\mutagen: Reliable Coverage-Guided, Property-Based Testing using Exhaustive Well-Typed Mutations

### 2021-04-01 papers/oopsla-21/main.tex
WIP

\usepackage{wrapfig}
\usepackage{xcolor}
\usepackage{color}
\usepackage{multicol}
\usepackage[ruled,vlined,noend]{algorithm2e}
\usepackage{xspace}
\usepackage{tikz}
\usepackage{tikzscale}
\usepackage{pgfplots}
\usepackage{paralist}
\pgfplotsset{compat=1.3}

@@

In both case studies, we additionally show how the heuristics described in
Section \ref{sec:heuristics} affect the testing performance by disabling them
when using our tool.

@@

Aliquam erat volutpat \citep{devai2013edsl}. Nunc eleifend leo vitae magna. In
id erat non orci commodo lobortis. Proin neque massa, cursus ut, gravida ut,
lobortis eget, lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus.
Mauris ac felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque
nec dui dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a,
aliquet quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet
tortor. Nam euismod tellus id erat.

@@

>>>>

\usepackage{wrapfig}
\usepackage{xcolor}
\usepackage{color}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{caption}
\usepackage[ruled,vlined,noend]{algorithm2e}
\usepackage{xspace}
\usepackage{tikz}
\usepackage{tikzscale}
\usepackage{pgfplots}
\usepackage{paralist}
\pgfplotsset{compat=1.3}

@@

In both case studies, we additionally show how the FIFO scheduling and trace
reset heuristics described in Section \ref{sec:heuristics} affect the testing
performance by disabling them when using our tool.
%
We call these variants \textit{no-FIFO} and \textit{no-reset}, respectively.
%
In the case of \textit{no-reset}, the amount of random mutations is no longer
controlled by this heuristic, so we fixed it to 25 random mutations throughout
the execution of the tool.

@@

The results of this case study are shown in Fig. \ref{fig:results:ifc}, ordered
by the failure rate achieved by \fuzzchick in decreasing order --- notice the
logarithmic scale used on the MTTF.
%
As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
with a \%100 success rate within the one hour testing budget.
%
In turn, our tool manages to find every bug on all runs, and taking less than a
minute in the worst absolute case.
%
These results are somewhat pesimistic towards \fuzzchick if compared to the ones
presented originally by \citeauthor{lampropoulos2019coverage}.
%
However, we want to remark that our experiments encompass 30 independent runs
instead of the 5 originally used when presenting \fuzzchick.
%
Moreover, since the MTTF simply aggregates all runs, regardless of if they found
a bug or timed out, this metric is quite sensitive to the particular timeout
used on each experiment, and will tend to inflate the results as soon as the
failure rate goes below 1.
%
For this reason, additionally comparing the failure rate between tools gives us
a better estimation of the reliability of both tools, where \mutagen shows a
clear improvement.

\begin{wrapfigure}{r}{0.5\textwidth}
\vspace{-10pt}
\footnotesize
\begin{tabular}{ccccc}
\hline
\multicolumn{1}{c}{
  \multirow{2}{*}{\textbf{Bug}}} &
  \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\fuzzchick}}} &
  \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\mutagen}}} &
  \multicolumn{2}{c}{\textbf{VD.A}} \\
\cline{4-5}
\multicolumn{1}{c}{} &
\multicolumn{1}{c}{} &
\multicolumn{1}{c}{} &
\multicolumn{1}{c}{\textbf{Value}} &
\multicolumn{1}{c}{\textbf{Estimate}} \\
\hline
1  & 13974.9 & 3632.7 & 0.89 & {\color{red} large} \\ \hline
2  & 23962.9 & 7122.3 & 0.90 & {\color{red} large} \\ \hline
3  & 19678.5 & 4633.9 & 0.89 & {\color{red} large} \\ \hline
4  & 20398.4 & 16831.2 & 0.72 & {\color{red} medium} \\ \hline
5  & 17348.1 & 2326.6 & 0.94 & {\color{red} large} \\ \hline
6  & 10727.1 & 2312.9 & 0.92 & {\color{red} large} \\ \hline
7  & 5070.7 & 332.9 & 0.91 & {\color{red} large} \\ \hline
8  & 5596.4 & 298.7 & 0.90 & {\color{red} large} \\ \hline
9  & 16402.4 & 26342.5 & 0.51 & negligible \\ \hline
10 & 11553.0 & 25044.5 & 0.55 & negligible \\ \hline
11 & 11304.3 & 2536.8 & 0.82 & {\color{red} large} \\ \hline
12 & 18507.3 & 14482.7 & 0.65 & {\color{red} small} \\ \hline
13 & 17961.5 & 13454.9 & 0.65 & {\color{red} small} \\ \hline
14 & 10621.9 & 3928.3 & 0.78 & {\color{red} large} \\ \hline
15 & 23866.3 & 43678.2 & 0.23 & {\color{blue} large} \\ \hline
16 & 25321.7  & 27602.0 & 0.54 & negligible \\ \hline
17 & 28515.6 & 52344.9 & 0.47 & negligible \\ \hline
18 & 24218.6 & 123401.3 & 0.14 & {\color{blue} large} \\ \hline
19 & 18638.9 & 30682.3 & 0.45 & negligible \\ \hline
20 & 18883.1 & 15841.9 & 0.60 & {\color{red} small} \\ \hline
& & \textbf{Mean} & 0.67 & {\color{red} medium} \\
\hline
\end{tabular}
\captionof{table}{\label{table:ifc:tests} Mean tests until first failure and effect size
  comparison between \fuzzchick and \mutagen when considering only succesfull
  runs.
  %
  Estimates in {\color{blue} blue} and {\color{red} red} indicate results
  favouring \fuzzchick and \mutagen, respectively. }
\vspace{-15pt}
\end{wrapfigure}

If we only consider the succesfull runs where \fuzzchick does not time out, we
observed a peculiarity.
%
This tool either finds bugs relatively quickly (after a few thouthands tests) or
does not find them at all within the time budget, which suggest that its power
scheduler assigns some mutation candidates too little energy before they become
uniteresting and get ultimately discarded.
%
Under this consideration, it is fair to think that \fuzzchick might be a better
choice if we set a shorter timeout and run it several times.
%
Thus, to be an improvement over \fuzzchick, our tool must be able to find bugs
not only more reliably, but also relatively fast!

Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
required by both tools to find the first failure when we only consider
successful runs.
%
Additionally, we compute the non-parametric Vargha-Delaney A measure \cite{vda,
  arcuri}, that encodes the probability of \mutagen to yield better results than
\fuzzchick.
%
As it can be observed, using an exhaustive mutation approach can be not only
more reliable, but also faster, where \mutagen is likely to find bugs faster
that \fuzzchick in 14 of the 20 bugs injected into the abstract machine.

@@

In terms of the \mutagen heuristics, we can also arrive to some conclusions.
%
In first place, this case study does not seem to be affected by using a FIFO
scheduling.
%
Upon inspection, we found that the reason behind this is that the mutation
candidate queues remain empty most of the time (generation mode), and when a new
interesting candidate gets inserted, all its mutants (and their descendents) are
processed before the next one is enqueued.

On the other hand, disabling the trace saturation heuristic (\textit{no-reset})
we observed two interesting effects:
%
havig fixed the amount of random mutations to 25 adds a seemingly constant
overhead when finding most of the (easier) bugs, suggesting that we are spending
worthless time mutating numeric subexpressions inside the generated machine
states and that a smaller number of random mutations could be equally effective
to find such bugs.
%
However, some of the hardest to find bugs cannot be reliably exposed using this
amount of random mutations (\#4, \#9, \#12, \#18 and \#10), suggesting that one
should increase this number even further to be able to discover all bugs within
the time budget.
%
In consequence, we believe this heuristic is effective at automatically tunning
this internal parameter of our tool, especially when the user is unsure about
what the best value for it might be.
